---
title: "'BOLTSSIRR' Package for BOLT-SSI: Fully Screening Interaction Effects for Ultra-High Dimensional Data "
output:
  pdf_document: 
    keep_tex: yes
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1 Overview

This vignette provides an introduction to the `BOLTSSIRR' package. BOLTSSI is a statistical approach for detecting interaction effects among predict variables to response variables, which is  often an crucial
step in regression modeling of real data for various applications.  Through this publicly available package, we provide a unified environment to carry out interaction pursuit using a simple sure screening procedure (SSI) to fully detect significant pure interactions between predict variables and the response variable in the high or ultra-high dimensional generalized linear regression models. Furthermore, we suggest to discretize continuous predict variables, and utilize the Boolean operation for the marginal likelihood estimates. The so-called 'BOLTSSI'  procedure is proposed to accelerate the sure screening speed of the procedure.

After screening the interaction effects, we just use penalzied likelihood function with LASSO penalty to further select the variables including the interaction terms. The objective function is 
$$-loglik/nobs+\lambda*penalty.$$


This vignette is organized as follows. Section 2 introduces the basic principle of our methods. Section 3 illustrates  how to screen the inteaction effecs and choose  the variables by using this package.


#2 Introduction
##2.1 SSI
Assume that given the predictor vector $x$, the conditional distribution of the random variable $Y$ belongs to an exponential family, whose  probability density function has the canonical form
$$f_{Y|x}(y|x)=\exp\{y\theta(x)-b(\theta(x))+c(y)\}$$
where $b(\cdot)$ and $c(\cdot)$ are some known functions and $\theta(x)$ is a canonical natural parameter. Here we ignore the dispersion parameter $\phi$ in the above density function, since we only concentrate on the estimation of mean regression function. It is well known that the distributions in the exponential family include the Binomial, Gaussian, Gamma, Inverse-Gaussian and Poisson distributions.

We consider the following generalized linear model with two-way interaction:
$$E(Y|X=x)=b'(\theta(x))=g^{-1}\left(\beta_0+\sum_{i=1}^p\beta_iX_i+\sum_{i<j}\beta_{ij}X_iX_j\right)
$$
for some link function $g(\cdot)$. And we focus on the canonical link function, hence $g^{-1}(\cdot)=b'$ and $$\theta(x)=\beta_0+\sum_{i=1}^p\beta_iX_i+\sum_{i<j}\beta_{ij}X_iX_j\triangleq\beta_0+\sum_{i=1}^p\beta_iX_i+\sum_{i<j}\beta_{ij}X_{ij}.$$

Our method is based on the marginal likelihood ratio screening, which builds on the difference between two marginal negative log-likelihood functions in the following two models:
$$E(Y|X_i,X_j)=\beta_{i,j0}+\beta_{i,}X_i+\beta_{j,}X_j$$
and 
$$E(Y|X_i,X_j)=\beta_{i,j0}+\beta_{i,}X_i+\beta_{j,}X_j+\beta_{ij}X_{ij}.$$
To ease the presentation, through this vignette, $X_{ij}$ is referred to as an important interaction if its regression coefficient $\beta_{ij}$ is nonzero. $X_i$ is called an active interaction variable if there exists some $1 \leq i\neq j \leq p$ such that $X_{ij}$ is an important interaction. Our method ``SSI'' is based on the above difference and to implemnet the screening procedure for interaction terms.

##2.2 BOLTSSI
Furthermore, the ``BOLTSSI''  procedure is proposed to accelerate the sure screening speed of the procedure.  The first step of BOLT-SSI is to discretize one continuous attribute by creating one categorical variabe wtih a specified number of levels. After discretization, the Boolean operation can be used to speedup SSI procedure, especially the algorithm to calculate the difference in the above section 2.1. All details can be seen in the paper (  et al. [2018]).






#3 Screening and Prediction


The package can be loaded with the command:
```{r eval=FALSE}
library(BOLTSSIRR)
```


##3.1 The function 'BOLT_SSI'

###Description
This function  implements the Sure Independence Screening SSI and BOLTSSI for interaction screening.

###Usage

```{r eval=FALSE}
BOLT_SSI(X,y,extra_pairs,code_num,thread_num)
```

###Arguments


Arguments    |
-------------|-------------------------------------------------------------------------------------------------
X            |The design matrix, of dimensions $n*p$, without an intercept. Each row is an observation vector.
y            |The response vector of dimension $n*1$.   
extra_pairs  |The number of remaining interaction effects by using SSI or BOLTSSI,default is $n$.
code_num     |The level of preditor varibales after discretization, default is 3.
thread_num   |The number of thread_num, default is 4.
             



###Value
Returns a matrix with three columns. 

Values                 |
-----------------------|---------------------------------------------------------------------------------------------
Th frist column        |The value $i$, where $i$ is the index of the avtive interaction variable $X_i$.
The second column      |The value $j$, where $j$ is the index of the avtive interaction variable $X_j$.  
The third column       |The  difference between two marginal negative log-likelihood functions.



###Examples

```{r eval=FALSE}
library(BOLTSSIRR)
set.seed(0)
p=300;n=100;rho=0.5
H<-abs(outer(1:p,1:p,"-"))
covxx=rho^H
cholcov = chol(covxx)
x0 = matrix(rnorm(n*p), n, p)
x = x0%*%cholcov

#gaussian response
set.seed(0)
y=2*x[,1]+2*x[,8]+3*x[,1]*x[,8]+rnorm(n)
model1=BOLT_SSI(x,y)
head(model1)

#binary response
set.seed(40)
feta = 2*x[,1]+2*x[,8]+3*x[,1]*x[,8]; 
fprob = exp(feta)/(1+exp(feta))
y = rbinom(n, 1, fprob)
model2=BOLT_SSI(x,y)
head(model2)
```


##3.2 The function 'CV_BOLT_SSI_RR'

###Description
This function  fits a generalized linear model via penalized maximum likelihood. The regularization path is computed for the lasso penalty at a grid of values for the regularization parameter lambda.  And it implements Cross-Validation for BOLTSSI.  It can deal with all shapes of data, including very large sparse data matrices.

###Usage

```{r eval=FALSE}
CV_BOLT_SSI_RR(X,y,extra_pairs,cod_num,nfold,nLambda,thread_num)
```


###Arguments


Arguments    |
-------------|-------------------------------------------------------------------------------------------------
X            |The design matrix, of dimensions $n*p$, without an intercept. Each row is an observation vector.
y            |The response vector of dimension $n*1$.   
extra_pairs  |The number of remaining interaction effects by using SSI or BOLTSSI,default is $n$.
code_num     |The level of preditor varibales after discretization, default is 3.
nfold        |The number of folds, default is 10.
nLambda      |The number of lambda, default is 100.
thread_num   |The number of thread_num, default is 4.



###Value
An object of class "CV_BOLTSSI_RR" is returned, which is a list with the ingredients of the crossvalidation
fit.

Values                 |
-----------------------|---------------------------------------------------------------------------------------------
lambdas                |The values of lambda used in the fits.
beta                   | A nvars x length(lambda) matrix of coefficients, stored in sparse column format ("CsparseMatrix").
lambda_min             |The  value of lambda when predicition error arrives at its minimum by Cross-Validation
index_min              |The index of optimal value
covs                   |The intercept vector of dimension $n*1$.
pairs                  |A vector of all indexes of interaction effects when lambda is condisered as it optimal value. 


###Examples
```{r eval=FALSE}
set.seed(0)
p=300;n=100
rho=0.5
H<-abs(outer(1:p,1:p,"-"))
covxx=rho^H
cholcov = chol(covxx)

x0 = matrix(rnorm(n*p), n, p)
x = x0%*%cholcov
#gaussian response
set.seed(0)
y=2*x[,1]+2*x[,8]+3*x[,1]*x[,8]+rnorm(n)
model3=CV_BOLT_SSI_RR(x,y,extra_pairs=p,nfold=5)


Lambdas=model3$lambdas
Beta=model3$beta
Lambda.min=model3$lambda_min
index.min=which(Lambdas==Lambda.min)
Pairs=t(matrix(model3$pairs,nrow = 2))

m=length(final_Beta)
main_index=which(final_Beta[1:p]!=0)
inter_index=which(final_Beta[(p+1):m]!=0)
final_inter_index=Pairs[inter_index,]
main_index
final_inter_index
```



##3.3 The function 'BOLT_Predict'
###Description
This function makes predictions from a cross-validated BOLTSSIRR model, using the  stored object, and the optimal value chosed for lambda.

###Usage
```{r eval=FALSE}
BOLT_Predict(x,fit)
```

###Arguments


Arguments    |
-------------|-------------------------------------------------------------------------------------------------
x            |Matrix of new values for $x$ at which predcitons ate to be made. Must be a matrix. 
fit          |Fitted ``CV_BOLT_SSI_RR'' object.   


###Value
Returns a predicted response vector . 



###Examples
```{r eval=FALSE}
newx=matrix(rnorm(30*p),30)
yhat=BOLT_Predict(newx,model3)
```



